---
title: "Project 2"
author: "Po Yi Liu"
date: "9/24/2021"
output: html_document
---

```{r, warning=FALSE, message=FALSE}
library(tidyverse)
library(tidymodels)
library(janitor)
library(skimr)
library(kableExtra)
library(GGally)
library(kableExtra) # -- make nice looking resutls when we knitt 
library(vip)        # --  tidymodels variable importance
library(fastshap)   # -- shapley values for variable importance 
library(rpart.plot) # -- plotting decision trees 
```


# import data
```{r}
organic <- read_csv("D:/fallclass/Predictive Analytics Data Mining/project2/organics.csv") %>% clean_names()
head(organic)
```
#check the null
```{r}
null_count <- function(c){
  sum(is.na(c))
}

res_00 <- organic %>%
  summarise(across(1:13,null_count)) %>% 
  pivot_longer(cols=1:13, names_to ="column", values_to="null_count") %>%
  mutate(null_pct = null_count / nrow(organic))

res_00%>%
  mutate(null_pct = round(null_pct,5))
```

#clean data
```{r}
organic%>%
  summarise(mean(dem_affl,na.rm=TRUE),median(dem_affl,na.rm=TRUE),
            mean(dem_age,na.rm=TRUE),median(dem_age,na.rm=TRUE),
            mean(prom_time,na.rm=TRUE),median(prom_time,na.rm=TRUE))
```
```{r}
organic%>%
  count(dem_gender)
organic%>%
  count(dem_cluster_group)
organic%>%
  count(dem_reg)
organic%>%
  count(dem_tv_reg)
```


```{r}
organic01<-organic%>%
  select(-dem_cluster,-target_amt)%>%
  mutate(dem_affl = if_else(is.na(dem_affl),(median(dem_affl,na.rm = TRUE)),dem_affl),
         dem_age = if_else(is.na(dem_age),(median(dem_age,na.rm = TRUE)),dem_age),
         prom_time = if_else(is.na(prom_time),(median(prom_time,na.rm = TRUE)),prom_time),
         dem_cluster_group = if_else(is.na(dem_cluster_group),"C",dem_cluster_group),
         dem_reg = if_else(is.na(dem_reg),"South East",dem_reg),
         dem_tv_reg = if_else(is.na(dem_tv_reg),"London",dem_tv_reg),
         dem_gender = if_else(is.na(dem_gender),"F",dem_gender),
         dem_gender = if_else(dem_gender=="U","M",dem_gender))
```




#check the null
```{r}
res_001 <- organic01 %>%
  summarise(across(1:11,null_count)) %>% 
  pivot_longer(cols=1:11, names_to ="column", values_to="null_count") %>%
  mutate(null_pct = null_count / nrow(organic01))

res_001%>%
  mutate(null_pct = round(null_pct,5))
```
```{r}
skim(organic01)
```


#convert numberic binary columns to factors
```{r}
organic01<-organic01%>%
  mutate(target_buy = as.factor(target_buy))
head(organic01)
```


## Explore target
```{r,warning=FALSE}
organic01 %>%
  group_by(target_buy) %>%
  summarize(n=n()) %>%
  ggplot(aes(x=target_buy, y=n)) +
  geom_col(stat="count", fill='#69b3a2') +
  geom_text(aes(label = n), vjust = 0)+
  labs(title = "How many customers purchase organic products?",
       x="0 = customer not purchase | 1 = customer purchase", y="numbers of customer purchase")

organic01 %>%
  group_by(target_buy) %>%
  summarize(n=n()) %>%
  ungroup() %>%
  mutate(pct = n/sum(n))
```

```{r}
explore_variables <- c("dem_affl","dem_age","prom_spend","prom_time")
categorical_variable <- c("dem_cluster_group","dem_gender","dem_reg","dem_tv_reg","prom_class")
```

#create function for chart
```{r}
relation_chart <- function(data,var,by){
  data%>%
  ggplot(aes(x={{by}}, y=target_buy)) +
  geom_boxplot() +
  labs(title = paste("Boxplot of target_buy and",explore_variables_name))
}

for (explore_variables_name in explore_variables) {
  organic01%>%relation_chart(by= .data[[explore_variables_name]])%>%print()
  
}
```

```{r}
relation_chart1 <- function(data,var,by){
  data%>%
  ggplot(aes(x={{by}})) + 
  geom_bar(aes(fill = target_buy), position = "fill")  +
  labs(title = paste("Histogram of target_buy and",categorical_variable_name))
}

for (categorical_variable_name in categorical_variable) {
  organic01%>%relation_chart1(by= .data[[categorical_variable_name]])%>%print()
  
}
```
# Build a decision tree 

## Convert character variables to factors (binary numerics were converted above)
```{r, warning=FALSE, message=FALSE}
organic01 %>%
    mutate_if(is.character, factor)  -> organic_prep

head(organic_prep)

```

## Partition your data 70/30 (train / test split) 

```{r}
set.seed(43)
organic_split <- initial_split(organic_prep, prop = 0.7)
organic_train <- training(organic_split)
organic_test <- testing(organic_split)
sprintf("Train PCT : %1.2f%%", nrow(organic_train)/ nrow(organic_prep) * 100)
sprintf("Test  PCT : %1.2f%%", nrow(organic_test)/ nrow(organic_prep) * 100)
head(organic_train)
```


## Recipe

```{r}
# -- create our recipe -- 
organic_recipe <- recipe(target_buy ~ ., data = organic_train) %>%
  step_dummy(all_nominal(), -all_outcomes()) %>%
    prep()

organic_recipe
```


## Bake 


```{r}
# -- apply the recipe 
bake_train <- bake(organic_recipe, new_data = organic_train)
bake_test  <- bake(organic_recipe, new_data = organic_test)
```

## Fit a logistic model

```{r}
logistic_glm <-logistic_reg(mode = "classification") %>%
                  set_engine("glm") %>%
                  fit(target_buy ~ ., data = bake_train)

tidy(logistic_glm) %>%
  mutate_at(c("estimate", "std.error", "statistic", "p.value"),round, 4)
```


## Stepwise logistic

```{r}
library(MASS)
steplog <- glm(target_buy ~ ., data = bake_train, family=binomial(link="logit"))
step <- stepAIC(steplog, direction="both")
summary(step)

```

```{r}
organic_steprecipe <- recipe(target_buy ~ dem_affl + dem_age + dem_cluster_group
                            + dem_gender , data = organic_train) %>%
  step_dummy(all_nominal(), -all_outcomes()) %>%
  prep()


bake_steptrain <- bake(organic_steprecipe, new_data = organic_train)
bake_steptest  <- bake(organic_steprecipe, new_data = organic_test)

logistic_step1 <-logistic_reg(mode = "classification") %>%
                  set_engine("glm") %>%
                  fit(target_buy ~ ., data = bake_steptrain)


tidy(logistic_step1) %>%
  mutate_at(c("estimate", "std.error", "statistic", "p.value"),round, 4)

```

## Fit a tree model 

```{r}
organic_tree <- decision_tree(mode="classification") %>%
                  set_engine("rpart") %>%
                  fit(target_buy ~ ., data = bake_train)


organic_tree$fit

options(scipen=0)
rpart.plot(organic_tree$fit, roundint=FALSE, extra = 3)

```

2possible
1. age<45, wealthly dem_affl>9.5
2.age>45, wealthy dem_affl>13, age<40, gender=female
### -  try a diff tree 
```{r}
organic_tree2 <- decision_tree(mode="classification",
                            cost_complexity = 0.001, #will change the potential for the tree to keep split
                            tree_depth = 5, #the tree depth will split 4 time if available
                            min_n = 3) %>%
                    #An integer for the minimum number of data points in a node that are required for the node to be split further
                  set_engine("rpart") %>%
                  fit(target_buy ~ ., data=bake_train)

organic_tree2$fit

options(scipen = 0) #draw scientific graph

rpart.plot(organic_tree2$fit, roundint=FALSE, extra=3)

```

target_buy ~ dem_affl + dem_age + dem_cluster_group+
                  dem_gender + dem_reg + dem_tv_reg + prom_class +
                  prom_spend + prom_time

#tuning the parameter
```{r}
colnames(organic_prep)

organic_prep[,!names(organic_prep)%in%"target_buy"]
```


```{r}
library(caret)
library(lattice)
set.seed(123)
treemod2 <- train(target_buy ~ dem_affl + dem_age + dem_cluster_group_C + 
                  dem_cluster_group_F + dem_gender_M + dem_tv_reg_Yorkshire,
                  data = bake_train,
                  method = "rpart",
                  trControl = trainControl("cv", number = 10),
                  tuneLength = 10)

plot(treemod2)
treemod2$bestTune
```

```{r}
organic_tree_bestfit <- decision_tree(mode="classification",
                            cost_complexity = 0.00156) %>%
                  set_engine("rpart") %>%
                  fit(target_buy ~ ., data=bake_train)

organic_tree_bestfit$fit
options(scipen = 0) 
rpart.plot(organic_tree_bestfit$fit, roundint=FALSE, extra=3)
```


## Prep for Evaluation (bank tree)


```{r}

# -- training 
predict(organic_tree, bake_train, type = "prob") %>%
  bind_cols(.,predict(organic_tree, bake_train)) %>%
  bind_cols(.,bake_train) -> scored_train_tree

head(scored_train_tree)

# -- testing 
predict(organic_tree, bake_test, type = "prob") %>%
  bind_cols(.,predict(organic_tree, bake_test)) %>%
  bind_cols(.,bake_test) -> scored_test_tree

head(scored_test_tree)
```


## Evaluate
```{r}
# -- AUC: Train and Test 
scored_train_tree %>% 
  metrics(target_buy, .pred_1, estimate = .pred_class) %>%
  mutate(part="training") %>%
  bind_rows( scored_test_tree %>% 
               metrics(target_buy, .pred_1, estimate = .pred_class) %>%
               mutate(part="testing") 
  ) 
  
# -- Variable Importance top 10 features  
organic_tree %>%
  vip(num_features = 5)
options(yardstick.event_first = FALSE)
# -- ROC Charts 
scored_train_tree %>%
  mutate(model = "train") %>%
  bind_rows(scored_test_tree %>%
              mutate(model="test")) %>%
  group_by(model) %>%
  roc_curve(target_buy, .pred_1) %>%
  autoplot()


# -- Confustion Matricies  
scored_train_tree %>%
  conf_mat(target_buy, .pred_class) %>%
  autoplot( type = "heatmap") +
  labs(title="Train Confusion Matrix")

scored_test_tree %>%
  conf_mat(target_buy, .pred_class) %>%
  autoplot( type = "heatmap") +
  labs(title="Test Confusion Matrix")

```

#organic tree 2
```{r}

# -- training 
predict(organic_tree2, bake_train, type = "prob") %>%
  bind_cols(.,predict(organic_tree2, bake_train)) %>%
  bind_cols(.,bake_train) -> scored_train_tree2

head(scored_train_tree2)

# -- testing 
predict(organic_tree2, bake_test, type = "prob") %>%
  bind_cols(.,predict(organic_tree2, bake_test)) %>%
  bind_cols(.,bake_test) -> scored_test_tree2

head(scored_test_tree2)
```

```{r}
# -- AUC: Train and Test 
scored_train_tree2 %>% 
  metrics(target_buy, .pred_1, estimate = .pred_class) %>%
  mutate(part="training") %>%
  bind_rows( scored_test_tree2 %>% 
               metrics(target_buy, .pred_1, estimate = .pred_class) %>%
               mutate(part="testing") 
  ) 
  
# -- Variable Importance top 10 features  
organic_tree2 %>%
  vip(num_features = 5)
options(yardstick.event_first = FALSE)
# -- ROC Charts 
scored_train_tree2 %>%
  mutate(model = "train") %>%
  bind_rows(scored_test_tree2 %>%
              mutate(model="test")) %>%
  group_by(model) %>%
  roc_curve(target_buy, .pred_1) %>%
  autoplot()


# -- Confustion Matricies  
scored_train_tree2 %>%
  conf_mat(target_buy, .pred_class) %>%
  autoplot( type = "heatmap") +
  labs(title="Train Confusion Matrix")

scored_test_tree2 %>%
  conf_mat(target_buy, .pred_class) %>%
  autoplot( type = "heatmap") +
  labs(title="Test Confusion Matrix")

```
#organic tree best fit
```{r}

# -- training 
predict(organic_tree_bestfit, bake_train, type = "prob") %>%
  bind_cols(.,predict(organic_tree_bestfit, bake_train)) %>%
  bind_cols(.,bake_train) -> scored_train_tree3
head(scored_train_tree3)
# -- testing 
predict(organic_tree_bestfit, bake_test, type = "prob") %>%
  bind_cols(.,predict(organic_tree_bestfit, bake_test)) %>%
  bind_cols(.,bake_test) -> scored_test_tree3
head(scored_test_tree3)
```

```{r}
# -- AUC: Train and Test 
scored_train_tree3 %>% 
  metrics(target_buy, .pred_1, estimate = .pred_class) %>%
  mutate(part="training") %>%
  bind_rows( scored_test_tree3 %>% 
               metrics(target_buy, .pred_1, estimate = .pred_class) %>%
               mutate(part="testing") 
  )

# -- Variable Importance top 10 features  
organic_tree_bestfit %>%
  vip(num_features = 5)
options(yardstick.event_first = FALSE)
# -- ROC Charts 

scored_train_tree3 %>%
  mutate(model = "train") %>%
  bind_rows(scored_test_tree3 %>%
              mutate(model="test")) %>%
  group_by(model) %>%
  roc_curve(target_buy, .pred_1) %>%
  autoplot()

# -- Confustion Matricies  
scored_train_tree3 %>%
  conf_mat(target_buy, .pred_class) %>%
  autoplot( type = "heatmap") +
  labs(title="Train Confusion Matrix")

scored_test_tree3 %>%
  conf_mat(target_buy, .pred_class) %>%
  autoplot( type = "heatmap") +
  labs(title="Test Confusion Matrix")

```

## Prep for Evaluation logistic_lm
```{r}
# -- training 
predict(logistic_step1, bake_steptrain, type = "prob") %>% #type prob convert 0,1 result to prob
  bind_cols(.,predict(logistic_step1, bake_steptrain)) %>% #no type=prob so result is 0, 1
  bind_cols(.,bake_steptrain) -> scored_train_glm

head(scored_train_glm)

# -- testing 
predict(logistic_step1, bake_steptest, type = "prob") %>%
  bind_cols(.,predict(logistic_step1, bake_steptest)) %>%
  bind_cols(.,bake_steptest) -> scored_test_glm

head(scored_test_glm)
```

## Evaluate


```{r}
# -- AUC: Train and Test 
scored_train_glm %>% 
  metrics(target_buy, .pred_1, estimate = .pred_class) %>%
  mutate(part="training") %>%
  bind_rows( scored_test_glm %>% 
               metrics(target_buy, .pred_1, estimate = .pred_class) %>%
               mutate(part="testing") 
  ) 
  

# -- Variable Importance top 10 features  
logistic_step1 %>%
  vip(num_features = 5)
options(yardstick.event_first = FALSE)
# -- ROC Charts 
scored_train_glm %>%
  mutate(model = "train") %>%
  bind_rows(scored_test_glm %>%
              mutate(model="test")) %>%
  group_by(model) %>%
  roc_curve(target_buy, .pred_1) %>%
  autoplot()



# -- Confustion Matricies  
scored_train_glm %>%
  conf_mat(target_buy, .pred_class) %>%
  autoplot( type = "heatmap") +
  labs(title="Train Confusion Matrix")

scored_test_glm %>%
  conf_mat(target_buy, .pred_class) %>%
  autoplot( type = "heatmap") +
  labs(title="Test Confusion Matrix")

```
```{r}
# -- training 
predict(logistic_glm, bake_train, type = "prob") %>% #type prob convert 0,1 result to prob
  bind_cols(.,predict(logistic_glm, bake_train)) %>% #no type=prob so result is 0, 1
  bind_cols(.,bake_train) -> scored_train_glm01

head(scored_train_glm01)

# -- testing 
predict(logistic_glm, bake_test, type = "prob") %>%
  bind_cols(.,predict(logistic_glm, bake_test)) %>%
  bind_cols(.,bake_test) -> scored_test_glm01

head(scored_test_glm01)

# -- AUC: Train and Test 
scored_train_glm01 %>% 
  metrics(target_buy, .pred_1, estimate = .pred_class) %>%
  mutate(part="training") %>%
  bind_rows( scored_train_glm01 %>% 
               metrics(target_buy, .pred_1, estimate = .pred_class) %>%
               mutate(part="testing") 
  ) 
options(yardstick.event_first = FALSE)
```






















